---
title: "Untitled"
output: html_document
date: "2024-10-26"
---
1.) (Use R) Consider the following data describing the sales of some items at a hardware
store. The variable names are the order number, the item, the region, the price and the
quantity, respectively. Input this data, and create a new variable called total_sales which is the product between price and quantity. Finally, print your new dataset.
```{r}
setwd("/Users/jacobrichards/Desktop/Stat_405_R/Stat 405 M10 HW")
data <- read.table("/Users/jacobrichards/Desktop/Stat_405_R/Stat 405 M10 HW/HW10.txt")

colnames(data) <- c("order","item","region","price","quantity")

price <- data[,4]
quantity <- data[,5]

total_sales <- price*quantity

final_dataset <- cbind(data,total_sales)

print(final_dataset)
```
2.) (Use R) Write a code to read in the dataset called “baseball.” The variables for each
observation, in order, are: first name, last name, height(feet-inches), weight(pounds),
position, number of years on team, birthdate, and date they joined the starting lineup (if
ever). Print the entire data set with the players sorted by position (alphabetically), and by
years on the team within each position (list most experienced players first).
```{r}

```




4.) (Use R) Consider the data set ‘exchange rates’ contains exchange rates from widely
traded rates of 22 different countries: US, Japan, Euro, and so on, on consecutive days.
Make a table of the average rate for each country per year.
```{r, ECHO=FALSE}
setwd("/Users/jacobrichards/Desktop/Stat_405_R/Stat 405 M10 HW")
mat <- as.data.frame(read.csv(file = "exchange rates.csv", header = TRUE))
d1 <- mat[,1]
d2 <- as.Date(d1, tryFormats = "%m/%d/%Y")
d3 <- format(d2, "%Y")
d4 <- as.numeric(d3)
mat[,1] <- d4
rownames(mat) <- seq(3779)

years <- seq(1995,2010,by=1)

means <- matrix(0,nrow=15,ncol=22)
     
for (i in 1:15) {
  year_data <- mat[mat[,1] == years[i], 2:23]  
  means[i, ] <- colMeans(year_data, na.rm = TRUE)  
}

year_means <- cbind(years,means)
colnames(year_means) <- colnames(mat)
print(year_means)
```
**undergraduate extra credit**
5.) (Use R) This problem is required for Grad students (those of you registered in
STAT 505) but can be extra credit for undergrad students. 
A researcher needs your expertise to combine the so called SRH (Health rate) before and
after a treatment. A random sample of size n = 2000 individuals were selected. A random
sample of 1000 individuals were then subjected to the treatment and another 1000 to the
control group. In the first data file called “SRH”, their status pre and post experiment are
reported in. The weights of the treatment and control group were then record in the files
called “weight_treatment” and “weight_control.” Combine the 3 data files properly to
create an analytic dataset as shown in the sample below:
Remember to sort the id in increasing order and sort the time in PRE and POST order.
Provide code for how you arrived at your analytic data set.
```{r}
SRH <- as.data.frame(read.csv(file = "/Users/jacobrichards/Desktop/Stat_405_R/Stat 405 M10 HW/SRH(1).csv", header = TRUE))
weight_control <- as.data.frame(read.csv(file = "/Users/jacobrichards/Desktop/Stat_405_R/Stat 405 M10 HW/weight_control(1).csv", header = TRUE))
weight_treatment <- as.data.frame(read.csv(file = "/Users/jacobrichards/Desktop/Stat_405_R/Stat 405 M10 HW/weight_treatment(1).csv", header = TRUE))
```

```{r}
#multiple input arguments of an order function will be mutually satisfied, where if a condtion is passed: the partiton indicated by a TRUE will be ordered in aescending. As both conditions of the order must be mutually satisfied, the ordering of id is preserved while Post is ordered in aescending (after pre).

SRH_id_sorted <- SRH[order(SRH[,1], SRH[,3] == "POST"), ]

WC_id_sorted <- weight_control[order(weight_control[,1]),]

WT_id_sorted <- weight_treatment[order(weight_treatment[,1]),]

(ID_Pre_WT <- na.omit(WT_id_sorted[,1:2]))

(ID_Post_WT <- na.omit(WT_id_sorted[,c(1,3)]))

ID_Pre_WC <- na.omit(WC_id_sorted[,1:2])

ID_Post_WC <- na.omit(WC_id_sorted[,c(1,3)])


Append <- numeric(nrow(SRH_id_sorted))

Final <- cbind(SRH_id_sorted,Append)


final[,4] <- 


# Sample data
data <- data.frame(
  id = c(1, 2, 3, 4, 5),
  group = c("A", "B", "A", "B", "A"),
  score = c(75, 90, 88, 85, 78)
)

# Sort by group ("A" first), then prioritize scores above 80
data_sorted <- data[order(data$group == "A", data$score > 80), ]
print(data_sorted)





```








